A Quantitative Genetic Linked Selection Model of the Negative Selection Process in Humans

### 2023-07-12 manuscript/manuscript.tex
fig shuffle

\begin{align}
    \label{eq:main_Ne}
    N_d = \frac{N}{Q^2 V_A + 1}
\end{align}
%
(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Section
\ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
forward-time model is that the inflation factor is invariant with respect to
the particular fitness background the neutral allele becomes stochastically
associated with. By contrast, the difficulty with modeling draft backwards in
time is that the coalescence rates experienced by a lineage are not invariant
to which lineage was sampled due to its particular associated fitness
background.

@@

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^n V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}

where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

\begin{align}
    N_d = N \exp \left( - \sum_i^n \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}

@@

\begin{align}
  \label{eq:main_eqns}
  N_d &= N \exp \left( -V_a \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
\end{align}

where $V_a = (U-R)s$ as in Equation \eqref{eq:Va}. We denote the solutions to
these equation, which represent equilibria under mutation-selection-drift-draft
process, as $\widetilde{N}_d$ and $\widetilde{R}$. These equilibria also imply
an equilibrium level of additive fitness variation $\widetilde{V}_a$ in the
segment, are used to calculate the reduction at an arbitrary position in the
genome (see Methods

With the equilibrium $V_a$ calculated at each segment for a particular $\mu$
and $s$, we can estimate the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any
genomic position $x$ (see Methods XXX).

\section*{Results: Validation of Theory}

Similar to previous work, our method characterizes the negative selection
process through the signal left on nucleotide diversity at linked sites.
Although nucleotide diversity is a less-informative statistic than the full
site frequency spectrum, it is much easier to derive models of how selection
alters expected pairwise coalescence times at linked sites than to describe the
entire genealogy. Spatial patterns of pairwise diversity are highly variable
across eukaryotic genomes, as linked selection leads diversity to become
correlated with spatial heterogeneity in recombination rates and the density of
conserved sites. Statistical methods then fit observed diversity to the
expected level under a particular theoretic model of linked selection
processes. Consequently, statistical inferences about selection may be
incorrect when the underlying theory inaccurately predicts average pairwise
coalescence times.

\subsection*{Simulations of a Segment under Negative Selection}

@@

Given that Santiago and Caballero's (\citeyear{Santiago2016-mu}) model is a
haploid-model for a single region under equilibrium negative selection, we
first sought to confirm it could accurately predict the reduction in realistic
forward-time simulations of diploid populations with recombination. Using SLiM
\parencite{Haller2019-vu,Haller2023-uk} we simulated a region of $10^5$
basepairs under varying levels of mutation and selection (see Methods
\ref{sec:methods-sim}). We find a close correspondence between the observed
and predicted reductions in effective population size $B=\nicefrac{N_d}{N}$
over all selection and mutation parameters including weak selection (Figure
\ref{fig:figure-1}A), in contrast to classic BGS theory. Furthermore, to
investigate whether this accuracy was caused by the model correctly predicting
the equilibrium fitness variance and ratchet rate, we also measured these
throughout the simulation. Again, we find the diploid SC16 theory accurately
predicts both the ratchet rate (Figure 1\ref{fig:figure-1}B) and the genic
fitness variance (Figure 1\ref{fig:figure-1}C).

@@

\subsection*{Chromosome-wide Simulations and Models of Negative Selection}

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general mutation-selection processes,
we extended their model so that it could be applied to whole-genome data. Our
software method \texttt{bgspy} numerically solves Equations XXX to compute the
equilibrium additive genic fitness variance ($V_a$) and ratchet rate ($R$)
across grids of mutation rates and selection coefficients. This is done for
each pre-specified segment in the genome that may be under negative selection
(e.g. coding sequences or UTRs). Then, these equilibrium can be used to
calculate a reduction map $B(x)$ across genomic positions $x$. To distinguish
between McVicker's B maps based on classic BGS theory, we call our reduction
map $B(x)$ a B' map.

To assess our predicted B' reduction maps, we simulated negative selection for
fixed mutation rates and selection coefficient on human chromosome 10 using
realistic recombination maps and conserved feature positions. Averaging over
$100$ replicates, we estimated the simulation expected reduction map and
compared this to the predicted B and B' maps for the same parameters. We find
that our B' maps and the classic BGS theory B maps closely match simulations
when selection is strong (top row of Figure \ref{fig:figure-1}E). This provides
the first realistic forward-time chromosome-scale simulation confirmation of
classic BGS theory. However, we find slight discrepancies in regions with low
recombination (Figure \ref{fig:figure-1}E). Second, we find our theory is
vastly more accurate than the classic B maps when selection is very weak ($2N_e
s \ll 1$; bottom row of Figure \ref{fig:figure-1}E), which is expected since
BGS theory does not work in this domain. Across all mutation and selection
parameters simulated, the relative error of the classic B maps is 14.6\%
whereas the relative error in the new B' maps is 5\%. Additionally, we find
that the mean squared error between simulations and B' maps is close to the
theoretic lower bound set by the coalescence variance \parencite[see Methods
XXX]{Tajima1983-gu}.

Nearly all of the error between the new B' maps and chromosome-wide simulations
occurs around the drift-barrier domain of $2Ns \approx 1$ (Figure
\ref{fig:figure-1}D). The higher error in this domain occurs across all
mutation rates. We hypothesized that the higher error in this domain may be
because when our method numerically solves Equations \eqref{eq:main_eqns} for
each segment, it does not consider the reduction experienced due to selection
at other linked segments. In particular, we use a fixed drift-effective
$N=1000$ corresponding to the number of diploids in the simulations rather than
$B(x) N$ at position $x$. To test this, we implemented a locally-rescaled
version of the B' maps, which numerically solves Equations \eqref{eq:main_eqns}
for each segment at position $x$ taking into account the reduction experienced
due to selection at other segments by setting $B'(x)N$ as the population size
(see Method XXX). We find the locally-rescaled B' maps reduce the relative
error from 5\% to 0.4\% and mean squared error (Figure \ref{fig:figure-1}D,
dashed colored lines), but does not entirely eliminate the error in the $2Ns
\approx 1$ domain. Overall, while we find evidence that local-rescaling
slightly reduces error in the $2Ns \approx 1$ domain, we do not use it during
in our genome-wide inference method. This is because to avoid circularity and
possible overfitting, the method would need to solve the equilibria equations
for all segments simultaneously, but this is currently computational
unfeasible.

@@

@@

Our composite likelihood method takes tracks of annotated features (an
``annotation model") that are \emph{a priori} expected to have similar fitness
effects, and estimates the overall mutation rate and distribution of fitness
effects for each feature type. We consider two classes of annotation models:
(1) CADD-based models, which consider the top $x\%$ of basepairs according to
the CADD pathogenicity score assigned to each basepair, and (2) and more
interpretable gene feature-based models that includes protein coding regions,
intron and UTRs, and PhastCons regions. We include PhastCons regions because
there is strong evidence of highly-conserved non-coding regions
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, that
would be missed by gene feature only annotation. These two classes of
annotation models have a trade-off between fine-scaled specificity to which
basepairs are likely to be under negative selection, and interpretability of
the DFE estimates for each feature.

Our method estimates the distribution of fitness effects (DFE) for each feature
class. While CADD-based models only have a single conserved feature class (e.g.
CADD 6\%), feature-based models can have multiple feature classes under varying
levels of selective constrain. However, overlapping features (e.g. a basepair
that is annotated as both PhastCons and coding sequence) must be assigned to
one category or the other. Since this assignment impact DFE estimates, we fit
both of the two alternatives. First, a \emph{PhastCons Priority} model, where
genic features that overlap PhastCons regions are classified as PhastCons, and
all remaining coding basepairs are labeled as CDS. Second, a \emph{Feature
Priority} model, where all coding basepairs are assigned to CDS, and the
PhastCons class catches the remaining highly-conserved non-genic regions.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all 
    populations (colors) fit at the megabase-scale. Round points are our B' 
    method and diamonds are the classic B. Lighter color round points are
    the out-sample $R_\text{LOCO}^2$ estimates for our B' method, and arrows show the decline
    in goodness-of-fit due to in-sample overfitting (out-sample $R_\text{LOCO}^2$ 
    were not calculated for classic B values due to computational costs). The 
    horizontal dashed lines are the $R_\text{drift}^2$ expected when the residual
    variance is given by the theoretic variance in coalescence times due to drift alone.}
  \label{fig:figure-2}
\end{figure}

Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
as well (Figure \ref{fig:figure-2}A), consistent with recent work using classic
BGS theory \parencite{Murphy2022-sj}. However, we find that our models predict
out-sample diversity levels slightly better than previous methods. For these
two models, we find that our B' method predicts $R_\text{LOCO}^2=68.45$\% and
$R_\text{LOCO}^2=68.04$\% of the out-sample variance in Yoruba pairwise
diversity at the megabase-scale, respectively. By contrast, the best-fitting
CADD 6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in
left-out 2 Mbp windows across YRI samples. We note that this might be explained
by other differences in data process, reference genome, etc.

In populations impacted by the out-of-Africa bottleneck, the goodness-of-fit
was lower across all models (e.g. 62.98\% and 59.98\% for CEU and CHB
respectively in the PhastCons Priority model). Our PhastCons Priority model
fits marginally better than the CADD 6\% model consistently across all three
populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
differences among the four feature types. By contrast, the CADD 6\% track must
fit patterns of diversity to conserved sites under a single DFE. The CADD 8\%
model fit slightly worse than the CADD 6\% across all populations, and the
Feature Priority model had the poorest-fit.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_3.pdf} 
    \caption{The distribution of fitness effects of new mutations estimates for
        Yoruba individuals. (A) The DFEs using sparse (left column) and
        full-coverage (right column) tracks, across different annotation models
        (row). Color indicates the feature type. (B) The DFE of the
        full-coverage CDS-priority model comparing the estimates across
    populations.}
  \label{fig:figure-3}
\end{figure}

Our composite likelihood method has three sets of parameters: expected
diversity in the absence of linked selection diversity $\pi_0$, the mutation
rate $\mu$, and the distribution of fitness effects $\mathbf{W}$ across the
selection grid for each of the $K$ feature types. Given that the relationship
between $\pi$ and the strength selection is U-shaped, our model in theory could
fit under different combinations of parameters.

However, across all of our annotation models, new deleterious mutations in
conserved feature classes were consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
pathogenic CADD scores. However, we find that there is considerable uncertainty
in how strong selection is against the most deleterious mutations. For example,
in the CADD 6\% full-track model, new mutations are estimated to have a
selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
the other CADD models. Moreover, these DFE estimates suggest a very high
average selection coefficient across CADD 6\% regions of ${\bar{s}}
\approx 0.043$ for YRI.

@@

\begin{figure}[htbp] 
    \centering
    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
    \caption{(A) }
  \label{fig:figure-4}
\end{figure}

@@

\subsection*{Tests of Parameter Inference using Forward Simulations}

Given these the sensitivity of mutation rate estimates and the strong-selection
non-identifiability issue, we next sought to assess the estimation overall
accuracy of our composite likelihood method using forward simulations. In order
to have enough megabase-scale data to apply our method, we independently
evolved the first five human chromosomes and combined them into a ``synthetic"
genome. Due to the immense computational expense of conducting forward
simulations at this scale over numerous parameter combinations, we used fixed
mutation and selection parameters in putatively conserved regions (see Methods
XXX). We note three important findings from these simulations.

First, the linked selection signal, as measured by the $R^2$ between predicted
and simulated megabase-scale diversity, increases with the intensity of
selection against new deleterious mutations and mutation rate (Figure
\ref{fig:figure-4}B, top row). This shows the ``strength" of the linked
selection signal varies as expected with these key parameters. However, if new
mutations are highly deleterious ($s=0.05$), $R^2$ is reduced. This is
consistent with very strong selection having less localized effects and acting
more like a genome-wide faster rate of drift.

Second, both classic BGS theory and our B' method accurately infer the average
selection coefficient under strong selection (Figure \ref{fig:figure-4}B,
middle row). However, when selection was weak, the classic BGS model
erroneously estimates strong selection and a very low mutation rate. By
contrast, our B' method estimated selection coefficients much closer to their
true value. A minor discrepancy occurs around $2Ns = 1$, for reasons discussed
in Section XXX. Because we only simulated fixed selection coefficients and five
chromosomes, we only assessed the accuracy of average selection coefficients
and not the DFE probabilities.

Third, with both our B' and the classic BGS models, we find mutation rate
estimates show slight biases in the strong-selection regime (Figure
\ref{fig:figure-3}B, bottom row). Still, our B' method leads to more accurate
mutation rate estimates than classic BGS theory across a variety of selection
coefficients. These biases in mutation rate estimates suggest that benchmarking
genome-wide models via their concordance with pedigree-based rate estimates may
not be suitable. When BGS is not acting, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-4}B right column), all
estimates are worse. This makes intuitive sense; the overall linked selection
signal is weaker relative to drift-based noise. However, we note that this is
an unlikely region of parameter space and this issue is readily diagnosable
from the low $R^2$ values.

Finally, we also used simulations to test the sensitivity of our method to
violations of two assumptions: constant population size and recessivity of
selective effects. Since we find little difference in parameter estimates among
YRI and bottlenecked out-of-Africa (CEU and CHB) samples, we tested whether a
population expansion would bias our method. We use an expansion rate and timing
based on those estimated for Europeans by \textcite{Gutenkunst2009-pg}, finding
they had little effect on parameter estimates (Supplementary Section
\ref{supp:sim-assum}). By contrast, we find our estimates are downwardly biased
when mutational effects are weak and recessive, but are accurate under
recessive strong selection. This occurs because the fixation probabilities of
recessive mutations are in reality much lower than under our model that assumes
additivity.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_5.pdf}

@@

@@

\begin{align}
  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
  V_a &= (U - 2 R) s_j & \text{\emph{fitness variance equation}} \\
  Q^2(m_i, s_j) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
  Z &= 1 - \frac{Us_j}{U-{M}} & \text{\emph{variance decay rate}}
\end{align}
%
and $U = m_i L$ and $M = r_\text{BP} L$ are the total mutation and
recombination rates in the segment. A detailed derivation of these equations
can be found in Supplementary Section \ref{supp:theory}. The recombination
rate in a segment is determined by a user-supplied recombination map.

>>>>

\begin{align}
    \label{eq:main_Ne}
    N_d = \frac{N}{Q^2 V_A + 1}
\end{align}
%
(c.f. \cite{Robertson1961-ho,Santiago1995-hx}; see Supplementary Materials Section
\ref{supp:theory} for a proof). The benefit of modeling draft with Robertson's
forward-time model is that the inflation factor is invariant with respect to
the particular fitness background the neutral allele becomes stochastically
associated with. By contrast, the difficulty with modeling draft backwards in
time is that the coalescence rates experienced by a lineage are not invariant
to which lineage was sampled due to its particular associated fitness
background.

@@

\begin{align}
    \label{eq:polygenic_Ne}
    N_d \approx N \exp\left(-\sum_{i=1}^S V_{A,i} \frac{Q_i^2}{2}\right)
\end{align}

where the factor of one-half comes from ignoring weak associations from
unlinked regions and chromosomes (see Supplementary Materials Section
\ref{supp:heritable-fitness}). In our genome-wide model, we consider the
summation in Equation \eqref{eq:polygenic_Ne} over non-overlapping segments $i
\in \{1, 2, \ldots, S\}$ each undergoing selection such that segment $i$
contributes additive fitness variance $V_{A,i}$ to the total additive genetic
fitness variance. Under equilibrium levels of fitness variation, the
autocorrelation function for a neutral allele associated with segment $i$ is
$C(t) = [(1-r_i)(1-\kappa_i)]^t$, where $r_i$ is the recombination fraction to
the segment and $\kappa_i$ is the rate that the associated fitness variance
decays due to selective dynamics. Then, the cumulative autocorrelation is,

@@

\begin{align}
    N_d = N \exp \left( - \sum_i^S \frac{\mu L_i}{s(1 + r_i(1-s)/s)^2} \right) 
\end{align}

@@

\begin{align}
  \label{eq:main_eqns}
  N_d &= N \exp \left( -(U-R)s \frac{Q^2}{2} \right) & \text{\emph{draft-effective population equation}} \\
  R &= \frac{4N_d U s}{\exp(4 N_d s)-1}  & \text{\emph{ratchet equation}}
\end{align}
%

We denote the solutions to these equation, which represent equilibria under
mutation-selection-drift-draft process, as $\widetilde{N}_d$ and
$\widetilde{R}$. These equilibria also imply an equilibrium level of additive
fitness variation $\widetilde{V}_a$ in the segment, which are used to calculate
the reduction factor $B(x) = \nicefrac{N_d}{N}$ at any genomic position $x$
(see Methods \ref{sec:methods-maps}).

\section*{Results: Simulation Validation of Theory and Methods}

Given that modeling the interplay of mutation, drift, and draft under both weak
and strong selection has proven to be difficult problem, we first sought to
verify the SC16 theory and our genome-wide implementation of it with a three
levels of simulations.

\subsection*{Simulations of a Segment under Negative Selection}
\label{sec:segment-sims}

@@

Our first set of forward simulations were to ensure that the SC16 model
adequately captures selective dynamics in a single $10^5$ basepair region
undergoing negative selection, across a variety of mutation rates and selection
coefficients (see Methods \ref{sec:methods-sim}). We find a close
correspondence between the observed and predicted reductions in effective
population size $B=\nicefrac{N_d}{N}$ over all selection and mutation
parameters including weak selection (Figure \ref{fig:figure-1}A), in contrast
to classic BGS theory. Furthermore, to investigate whether this accuracy was
caused by the model correctly predicting the equilibrium fitness variance and
ratchet rate, we also measured these throughout the simulation. Again, we find
the diploid SC16 theory accurately predicts both the ratchet rate (Figure
1\ref{fig:figure-1}B) and the genic fitness variance (Figure
1\ref{fig:figure-1}C).

@@

\subsection*{Chromosome-wide Simulations and Models of Negative Selection}
\label{sec:chrom-sims}

Given the accuracy of the SC16 model in predicting the reduction factor $B$ and
ratchet rate for a single segment under general mutation-selection processes,
we next extended their model so that it could be applied to whole-genome data.
Our software method \texttt{bgspy} numerically solves Equations
\eqref{eq:main_eqns} to compute the equilibrium additive genic fitness variance
($V_a$) and ratchet rate ($R$) across grids of mutation rates and selection
coefficients. This is done for each pre-specified segment in the genome that
may be under negative selection (e.g. coding sequences or UTRs). Then, these
equilibrium can be used to calculate a reduction map $B(x)$ across genomic
positions $x$. To distinguish between McVicker's B maps based on classic BGS
theory, we call our reduction map $B(x)$ a B' map.

We next validated our predicted B' reduction maps with realistic
chromosome-scale forward simulations of negative selection in conserved
regions. Averaging over $100$ replicates, we estimated the simulation empirical
reduction map and compared this to the predicted B and B' maps for the same
parameters. We find that our B' maps and the classic BGS theory B maps closely
match simulations when selection is strong (top row of Figure
\ref{fig:figure-1}E). This provides the first realistic forward-time
chromosome-scale simulation confirmation of classic BGS theory. However, we
find slight discrepancies in regions with low recombination (Figure
\ref{fig:figure-1}E). Second, we find our theory is vastly more accurate than
the classic B maps when selection is very weak ($2N_e s \ll 1$; bottom row of
Figure \ref{fig:figure-1}E), which is expected since BGS theory does not work
in this domain. Across all mutation and selection parameters simulated, the
relative error of the classic B maps is 14.6\% whereas the relative error in
the new B' maps is 5\%. Additionally, we find that the mean squared error
between simulations and B' maps is close to the theoretic lower bound set by
the coalescence variance \parencite[see Methods XXX]{Tajima1983-gu}.

The discrepancy between our theoretic B' maps and chromosome-wide simulations
that leads to the 5\% error occurs entirely around the drift-barrier domain of
$2Ns \approx 1$ (Figure \ref{fig:figure-1}D). The higher error in this domain
occurs across all mutation rates. We hypothesized this error may be due to
selective interference between segments that is not taken into account when we
numerically solve Equations \eqref{eq:main_eqns} for each segment
independently. In particular, we use a fixed drift-effective $N=1000$
corresponding to the number of diploids in the simulations rather than $B(x) N$
at position $x$. To test this, we implemented a locally-rescaled version of the
B' maps, which numerically solves Equations \eqref{eq:main_eqns} for each
segment at position $x$ taking into account the reduction experienced due to
selection at other segments by setting $B'(x)N$ as the population size. We find
the locally-rescaled B' maps reduce the relative error from 5\% to 0.4\% and
mean squared error (Figure \ref{fig:figure-1}D, dashed colored lines), but does
not entirely eliminate the error in the $2Ns \approx 1$ domain. Overall, while
we find evidence that local-rescaling slightly reduces error in the $2Ns
\approx 1$ domain, we do not use it during in our genome-wide inference method.
This is because to avoid circularity and possible overfitting, our method would
need to solve the equilibria equations for all segments simultaneously;
unfortunately, this is currently computational unfeasible.

@@

\subsection*{Validation of Composite-likelihood Method using Forward
Simulations}

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_2.pdf} 
    \caption{Comparison of parameter estimates using classic BGS theory (green
        lines) with our new B' method (blue lines) across both full and sparse
        track types (dark versus light hue), and different mutation rates
        (columns). In the first row, $R^2$ between predictions and observations
        increases with selection intensity. Both classic BGS and B' methods fit
        the data equally well with sparse tracks, though the estimated
        selection coefficients are inaccurate when selection is weak (second
        row). Mutation rate estimates (third row) are more accurately estimated
        by the B' method than classic BGS, but overall show slight biases.
        Overall, classic BGS methods fail as expected when full-coverage tracks
    are used, since it cannot accommodate weak selection and neutrality. }
  \label{fig:figure-2}
\end{figure}

We next developed a composite-likelihood method that estimates the distribution
of selection coefficients for each feature type, the mutation rate, and the
diversity in the absence of linked selection ($\pi_0$) by fitting the theoretic
reduction map to genome-wide diversity (see Methods
\ref{sec:methods-likelihood}). We validated that our method can accurately
estimate the parameters of negative selection by simulation a ``synthetic
genome" of the first five human chromosomes (see Methods
\ref{sec:methods-sim}). We note three findings from these simulations.

First, we find the $R^2$ between predicted and simulated megabase-scale
diversity serves as a measure of the strength of the linked selection signal in
genome-wide data. The $R^2$ increases with the intensity of selection against
new deleterious mutations and mutation rate (Figure \ref{fig:figure-2}, top
row). Under just drift or weak negative selection, the variance in diversity is
driven by unstructured coalescence noise along the genome and the predicted
reduction map $B(x)$ does not fit the data well. However, we note that under
very strong selection ($s=0.05$), $R^2$ is reduced. This is like due to very
strong selection having less localized effects and impacting overall
genome-wide diversity \parencite{Santiago1995-hx,Robertson1961-ho}.

Second, both our implementation of classic BGS theory and our B' method
accurately infer the average selection coefficient under strong selection
(Figure \ref{fig:figure-2}, middle row). However, when selection was weak, the
classic BGS model erroneously estimates strong selection and a very low
mutation rate. By contrast, our B' method estimated selection coefficients much
closer to their true value. A minor discrepancy occurs around $2Ns = 1$, likely
due to the sensitivity of mutations in this region to selective interference
(discussed in Section \ref{sec:chrom-sims}). Because we only simulated fixed
selection coefficients and five chromosomes, we only assessed the accuracy of
average selection coefficients and not the DFE probabilities.

Third, we find slight biases in mutation rated estimated by both our B' and the
classic BGS methods (Figure \ref{fig:figure-2}, bottom row). However, mutation
rate estimates based on our B' method are more accurate than classic BGS theory
across a range of selection coefficients. Overall, the bias in estimated
mutation rates suggests that benchmarking genome-wide negative selection models
based on their agreement with pedigree-based rate estimates may not be
appropriate. When BGS is not functioning, either due to weak selection or a low
rate of deleterious mutations (Figure \ref{fig:figure-2}, right column), all
estimates deteriorate. This is understandable, as the overall signal from
linked selection weakens relative to drift-based noise. We should note, though,
that this is an unlikely region of parameter space and this issue can be
readily diagnosed from the low $R^2$ values.

Finally, we tested the sensitivity of our method to violations of two
assumptions---constant population size and recessivity of selective
effects---through simulations. Using an expansion rate and timing as estimated
for Europeans by \textcite{Gutenkunst2009-pg}, we found that a population
expansion had minimal impact on parameter estimates (Supplementary Materials
Section \ref{supp:sim-assum}). We did not test the influence of population
bottlenecks, since parameter estimates of out-of-Africa bottlenecked
populations (CHB and CEU) did not differ much from YRI estimates (see below).
However, these simulations did find that selection coefficient estimates were
downwardly biased when deleterious mutations were recessive and weak.
Recessivity did not bias selection coefficient estimates when selection was
strong; this is expected since strongly deleterious mutations are predominantly
are sufficiently rare that they are not often found in a homozygote. Overall,
the sensitivity to recessivity is likely due to the ratchet rate being
overestimated when selection is weak and recessive mutations drift up to higher
frequency and experience stronger than expected selection in homozygotes.

@@

Our composite likelihood method takes tracks of annotated features (an
``annotation model") that are \emph{a priori} expected to have similar fitness
effects, and estimates the overall mutation rate and distribution of fitness
effects for each feature type. We consider two classes of annotation models:
(1) CADD-based models, which consider the top $x\%$ most pathogenic basepairs
according to the CADD score, and (2) and more interpretable gene feature-based
models that includes protein coding regions, intron and UTRs, and PhastCons
regions. We include PhastCons regions because there is strong evidence of
highly-conserved non-coding regions
\parencite{Meader2010-hm,Harmston2013-tt,Katzman2007-gq,Siepel2005-wh}, that
would be missed by gene feature only annotation. These two classes of
annotation models have a trade-off between fine-scaled specificity to which
basepairs are likely to be under negative selection, and interpretability of
the DFE estimates for each feature.

Our method estimates the distribution of fitness effects (DFE) for each feature
class. While CADD-based models only have a single conserved feature class (e.g.
CADD 6\%), feature-based models can have multiple feature classes under varying
levels of selective constrain. However, overlapping features (e.g. a basepair
that is annotated as both PhastCons and coding sequence) must be assigned to
one category or the other. Since this assignment impacts DFE estimates, we fit
both of the two alternative models. First, a \emph{PhastCons Priority} model,
where genic features that overlap PhastCons regions are classified as
PhastCons, and all remaining coding basepairs are labeled as CDS. Second, a
\emph{Feature Priority} model, where all coding basepairs are assigned to CDS,
and the PhastCons class catches the remaining highly-conserved non-genic
regions.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_3.pdf}

    \caption{The $R^2$ estimates for a sparse (A) and full (B) models, for all
        populations (colors) fit at the megabase-scale. Round points are our B'
        method and diamonds are the classic BGS (we exclude classic BGS in the
        full track subfigure, since these all fit very poorly). Lighter color
        round points are the out-sample $R_\text{LOCO}^2$ estimates for our B'
        method, and arrows show the decline in goodness-of-fit due to in-sample
        overfitting (out-sample $R_\text{LOCO}^2$ were not calculated for
        classic B values due to computational costs). The horizontal dashed
        lines are the $R_\text{drift}^2$ expected when the residual variance is
    given by the theoretic variance in coalescence times due to drift alone.}

  \label{fig:figure-2}
\end{figure}

Overall, we find the PhastCons Priority and CADD 6\% models fit about equally
as well (Figure \ref{fig:figure-2}A), consistent with recent work using classic
BGS theory \parencite{Murphy2022-sj}. However, we find that our models predict
out-sample diversity levels slightly better than previous methods. For these
two models, we find that our B' method predicts $R_\text{LOCO}^2=67.3$\% and
$R_\text{LOCO}^2=68.7$\% of the out-sample variance in Yoruba pairwise
diversity at the megabase-scale, respectively. By contrast, the best-fitting
CADD 6\% model from \textcite{Murphy2022-sj} explained 60\% of diversity in
left-out 2 Mbp windows across YRI samples. We note that this difference could
be explained by other differences in data processing, optimization, etc.

In populations impacted by the out-of-Africa bottleneck, the goodness-of-fit
was lower across all models (e.g. 60.4\% and 58.7\% for CEU and CHB
respectively in the CADD 6\% model).

Our PhastCons Priority model fits
marginally better than the CADD 6\% model consistently across all three
populations (Supplementary Table \ref{supp:tbl-r2}). This is likely because the
PhastCons Priority model has more DFE parameters, and thus can fit finer-scale
differences among the four feature types. By contrast, the CADD 6\% track must
fit patterns of diversity to conserved sites under a single DFE. The CADD 8\%
model fit slightly worse than the CADD 6\% across all populations, and the
Feature Priority model had the poorest-fit.

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_4.pdf} 
    \caption{The distribution of fitness effects of new mutations estimates for
        Yoruba individuals. (A) The DFEs using sparse (left column) and
        full-coverage (right column) tracks, across different annotation models
        (row). Color indicates the feature type. (B) The DFE of the
        full-coverage CDS-priority model comparing the estimates across
    populations.}
  \label{fig:figure-3}
\end{figure}

Our composite likelihood method has three sets of parameters: expected
diversity in the absence of linked selection diversity $\pi_0$, the mutation
rate $\mu$, and the matrix of distribution of fitness effects $\mathbf{W}$
across the selection grid for each of the $K$ feature types. Given that the
relationship between $\pi$ and the strength selection is U-shaped, we wondered
whether the linked selection signal in humans could fit under a weak selection
regime.

However, across all of our annotation models, new deleterious mutations in
conserved feature classes were consistently estimated to have strongly
deleterious effects (Figure \ref{fig:figure-3}A), consistent with previous work
\parencite{McVicker2009-ax,Murphy2022-sj}. This includes features such as
coding sequences, PhastCons regions, or the top 6\% of basepairs with the most
pathogenic CADD scores.

However, we find that there is considerable uncertainty
in how strong selection is against the most deleterious mutations. For example,
in the CADD 6\% full-track model, new mutations are estimated to have a
selection coefficients of $s = 10^{-1}$ with 42\% chance, $s=10^{-2}$ with
7.8\% chance, and $s=10^{-3}$ with 41\% chance. This creates a bimodal DFE
(Figure \ref{fig:figure-3}A, second row, left column), which is mirrored across
the other CADD models. Moreover, these DFE estimates suggest a very high
average selection coefficient across CADD 6\% regions of ${\bar{s}}
\approx 0.043$ for YRI.

@@

\begin{figure}[htbp] 
    \centering
    \includegraphics[width=\textwidth]{figures/figure_5.pdf} 
    \caption{(A) }
  \label{fig:figure-4}
\end{figure}

@@

@@

@@

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/figure_7.pdf}

@@

\begin{align}
  %V(x) &= U s - 2 {R}(x) s & \text{\emph{fitness variance equation}} \\
  V_a &= (U - 2 R) s_j & \text{\emph{fitness variance equation}} \\
  Q^2(m_i, s_j) &= \frac{2}{(1-Z)(2-(2-M)Z)} & \text{\emph{linkage inflation factor}} \\
  Z &= 1 - \frac{Us_j}{U-{M}} & \text{\emph{variance decay rate}}
\end{align}
%
and $U = m_i L$ and $M = r_\text{BP} L$ are the total mutation and
recombination rates in the segment. A detailed derivation of these equations
can be found in Supplementary Materials Section \ref{supp:theory}. The recombination
rate in a segment is determined by a user-supplied recombination map.
